---
title: "East Van Euchre Club Analytics (December 3, 2017)"
output:
  html_document:
    self_contained: no
---

## Input
Reads in a CSV file (currently [Euchre Analytics December 3rd @ Anza - MASTER DATA.csv](Euchre Analytics December 3rd @ Anza - MASTER DATA.csv)) with tournament statistics and outputs. Relies on the following fields (as named):
- "Name"
- "Opponent 1"
- "Round"
- "Score: Hand XX", where XX is a number (1-12)
- "Called: Hand XX", where XX is a number (1-12): Should be "x" or "v"

## Output
The code prints a table of by-person summary statistics and a series of graphs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load, echo=FALSE}
# Load and aggregate data
library(data.table)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(DT) # Sortable table

# Load data
setwd("~/github//euchre_analytics/")
raw_data <- fread("Euchre Analytics December 3rd @ Anza - MASTER DATA.csv")
# Rename variables to something a little more usable
score.vars.old <- grep("Score: Hand [0-9]+", names(raw_data), value=T)
score.vars <- gsub("Score: Hand ([0-9]+)?", "score_\\1", score.vars.old)
setnames(raw_data, score.vars.old, score.vars)
call.vars.old <- grep("Called: Hand [0-9]+", names(raw_data), value=T)
call.vars <- gsub("Called: Hand ([0-9]+)?", "hand_\\1", call.vars.old)
setnames(raw_data, call.vars.old, call.vars)
# Create variables to split out score, calls, loner calls, and loner wins by hand
raw_data[, paste0("called_", 1:length(call.vars)):=lapply(.SD, function(x) x=="v" | x == "x"), .SDcols=call.vars]
raw_data[, paste0("lonertry_", 1:length(call.vars)):=lapply(.SD, function(x) x=="x"), .SDcols=call.vars]
for (i in 1:length(call.vars)) {
  raw_data[, paste0("lonerwin_", i):=get(paste0("lonertry_", i))==TRUE & get(score.vars[i])==4]
}

# Match in data from opponents to get opponent score by round
raw_data[`Opponent 1`=="Nat", `Opponent 1`:="Natalie"] # Fix a name inconsistency
opp_data <- raw_data[, c("Name", "Round", score.vars), with=F]
setnames(opp_data, score.vars, paste0("opp_", score.vars))
raw_data2 <- merge(raw_data, opp_data, by.x=c("Opponent 1", "Round"), by.y=c("Name", "Round"))

# Aggregate for each person-round
raw_data2[, c("total_score"):=rowSums(.SD), .SDcols=score.vars]
raw_data2[, c("opp_score"):=rowSums(.SD), .SDcols=paste0("opp_", score.vars)]
raw_data2[, c("times_called"):=rowSums(.SD), .SDcols=paste0("called_", 1:length(call.vars))]
raw_data2[, c("times_lonertry"):=rowSums(.SD), .SDcols=paste0("lonertry_", 1:length(call.vars))]
raw_data2[, c("times_lonerwin"):=rowSums(.SD), .SDcols=paste0("lonerwin_", 1:length(call.vars))]

# Aggregate by person
agg.vars <- c("total_score", "opp_score", "times_called", "times_lonertry", "times_lonerwin")
agg_data <- raw_data2[, lapply(.SD, sum),
                      by=list(name=Name),
                      .SDcols=agg.vars]
```

### Per-person statistics

```{r analysis1, echo=FALSE}
# Export agg_data tabulation
agg.labs <- c("Total score", "Opponents' score", "Times called", "Loner attempts", "Loner wins")
setorder(agg_data, name)
datatable(agg_data,
          rownames=FALSE,
          colnames=c("Name", agg.labs),
          options=list(pageLength=nrow(agg_data)))
```

```{r weirdbug, echo=FALSE}
rm(.Random.seed) # Fixes a very weird bug related to datatable
set.seed(1)
```

### Distributions

```{r analysis2, echo=FALSE}
agg_data_long <- melt(agg_data, id="name")
agg_data_long[, variable:=factor(variable, # Generate nice labels
                                 levels=agg.vars,
                                 labels=agg.labs)]
ggplot(agg_data_long, aes(x=value)) + geom_histogram(bins=6) +
  facet_wrap(~variable, ncol=2, scales="free_x") +
  labs(x=NULL, y=NULL)
```


### Relationship between times called and total score

```{r analysis3, echo=FALSE}
ggplot(agg_data, aes(y=total_score)) +
  geom_label_repel(aes(x=times_called, label=name)) +
  geom_smooth(method=lm, aes(x=times_called)) +
  labs(x="Times called", y="Total score")
```
